|

### 2021-12-09 05-Chp-CPRD-Analysis.Rmd
Range of small fixes (randomised, ApoE, etc)

---
bibliography: bibliography/references.bib
csl: bibliography/nature.csl
output:
  bookdown::word_document2: 
      toc: false
      toc_depth: 3
      reference_docx: templates/word-styles-reference-01.docx
      number_sections: false 
  bookdown::pdf_document2:
    template: templates/brief_template.tex
    pandoc_args: !expr rmdfiltr::add_wordcount_filter()  
  bookdown::html_document2: default
documentclass: book
---

@@

In this chapter, I present an analysis of a large population-based electronic health record dataset to investigate the relationship between lipid-regulating agent (LRA) use and dementia outcomes. The analysis aims to address two important limitations of the current evidence base as identified by the systematic review presented in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading).

@@

Participants were ineligible for our cohort if they were less than 40 years of age (as these patients are less likely to be prescribed a LRA), had less than 12 months of "research quality" data, were simultaneously prescribed more than one lipid-regulating agent (due to the difficult of assigning these to a single exposure group), or were diagnosed with an outcome of interest before or on the date of the index event (i.e. had less than one full day of follow-up).

@@

In this analysis, missing data were handled using a multiple imputation approach.[@sterne2009] Variables with missing observations were identified for inclusion in the imputation model. Nominal variables with missing values were modelled using multinomial logistic regression, while continuous variables were modelled using linear regression. As per best practice, all variables used in the analytic model, including the outcome, were included in the imputation model.[@moons2006] Using the MICE (Multiple Imputation by Chained Equations) command in STATA16, 20 imputed datasets were created.

Missing data were only considered problematic for variables where a numerical test result was expected (e.g. BMI), or where a code existed for the absence of the condition (e.g. categorical smoking status). This approach was necessary, as absence of a code for other treatments or conditions (e.g. statin use or dementia) was assumed to indicate absence of the treatment/condition rather than being considered missing.[@wells2013strategies]

@@

The number of events, time-at-risk and crude rates for each drug class, tabulated by dementia outcome, are shown in Table \@ref(tab:followUp-table). A substantial majority (`r percentage.statins`) of participants prescribed a lipid-regulating agent were prescribed a statin. I excluded the "Ezetimibe and statins" (n=`r eze_sta_n`) and "Nicotinic acid groups" (n= `r nag_n`) classes from subsequent class-based subgroup analyses based on the extremely small number of participants in these groups. Note that the "Ezetimibe and statins" treatment group represent those prescribed a single treatment containing both ezetimibe and statins, rather than those where the two treatments were prescribed concurrently.

@@

For the composite all-cause dementia outcome, we found treatment with a lipid-regulating agent was associated with a slightly increased risk (`r anydem_text`), which lies between the associations for the Alzheimer and non-Alzheimer dementia outcomes as would be expected. There was also some evidence that fibrates were associated with increased risk of all-cause dementia (`r anydem_fib_text`).

@@

This analysis serves to illustrative the importance of the code lists chosen to define the outcomes of interest, particularly if they are used to define competing outcomes (e.g. AD vs non-AD dementia). This different codes used by Smeeth _et al._, in addition to an analytical approach that adjusted for covariates defined after the index date, may go some way to explaining why our analysis obtained different results despite the substantial overlap in the data sources used.

@@

A further limitation stems from the possibility of uncontrolled confounding due to genetic factors. The number of _ApoE_ $\epsilon4$ alleles represents the strongest genetic risk factor for Alzheimer's disease, but also substantially increases LDL cholesterol levels,[@bennet2007] potentially prompting treatment with a statin or other lipid regulating agent. I was unable to control for _ApoE_ genotype in this analysis as I did not have access to genetic data on participants. As a result, any protective association between LRA use and the Alzheimer's disease outcomes may be masked by residual negative confounding by ApoE.

>>>>

---
bibliography: bibliography/references.bib
csl: bibliography/nature.csl
output:
  bookdown::pdf_document2:
    template: templates/brief_template.tex
    pandoc_args: !expr rmdfiltr::add_wordcount_filter()  
  bookdown::word_document2: 
      toc: false
      toc_depth: 3
      reference_docx: templates/word-styles-reference-01.docx
      number_sections: false 
  bookdown::html_document2: default
documentclass: book
---

@@

In this chapter, I present an analysis of a large population-based electronic health record dataset to investigate the relationship between lipid-regulating agent (LRA) use and dementia outcomes. The analysis aims to address two important limitations of the current evidence base as identified by the systematic review presented in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading).

@@

Participants were ineligible for the cohort if they were less than 40 years of age (as these patients are less likely to be prescribed a LRA), had less than 12 months of "research quality" data, were simultaneously prescribed more than one lipid-regulating agent (due to the difficult of assigning these to a single exposure group), or were diagnosed with an outcome of interest before or on the date of the index event (i.e. had less than one full day of follow-up).

@@

In this analysis, missing data was handled using a multiple imputation approach.[@sterne2009] Variables with missing observations were identified for inclusion in the imputation model. Nominal variables with missing values were modelled using multinomial logistic regression, while continuous variables were modelled using linear regression. As per best practice, all variables used in the analytic model, including the outcome, were included in the imputation model.[@moons2006] Using the MICE (Multiple Imputation by Chained Equations) command in STATA16, 20 imputed datasets were created.

Missing data was only considered problematic for variables where a numerical test result was expected (e.g. BMI), or where a code existed for the absence of the condition (e.g. categorical smoking status). This approach was necessary, as absence of a code for other treatments or conditions (e.g. statin use or dementia) was assumed to indicate absence of the treatment/condition rather than being considered missing.[@wells2013strategies]

@@

The number of events, time-at-risk and crude rates for each drug class, tabulated by dementia outcome, are shown in Table \@ref(tab:followUp-table). A substantial majority (`r percentage.statins`) of participants prescribed a lipid-regulating agent were prescribed a statin. I excluded the "Ezetimibe and statins" (N =`r eze_sta_n`) and "Nicotinic acid groups" (N = `r nag_n`) classes from subsequent class-based subgroup analyses based on the extremely small number of participants in these groups. Note that the "Ezetimibe and statins" treatment group represent those prescribed a single treatment containing both ezetimibe and statins, rather than those where the two treatments were prescribed concurrently.

@@

For the composite all-cause dementia outcome, I found treatment with a lipid-regulating agent was associated with a slightly increased risk (`r anydem_text`), which lies between the associations for the Alzheimer and non-Alzheimer dementia outcomes as would be expected. There was also some evidence that fibrates were associated with increased risk of all-cause dementia (`r anydem_fib_text`).

@@

This analysis serves to illustrative the importance of the code lists chosen to define the outcomes of interest, particularly if they are used to define competing outcomes (e.g. AD vs non-AD dementia). This different codes used by Smeeth _et al._, in addition to an analytical approach that adjusted for covariates defined after the index date, may go some way to explaining why my analysis obtained different results despite the substantial overlap in the data sources used.

@@

A further limitation stems from the possibility of uncontrolled confounding due to genetic factors. The number of _Apo_$\mathcal{E}4$ alleles represents the strongest genetic risk factor for Alzheimer's disease, but also substantially increases LDL cholesterol levels,[@bennet2007] potentially prompting treatment with a statin or other lipid regulating agent. I was unable to control for _Apo_$\mathcal{E}$ genotype in this analysis as I did not have access to genetic data on participants. As a result, any protective association between LRA use and the Alzheimer's disease outcomes may be masked by residual negative confounding by _Apo_$\mathcal{E}$.

### 2021-12-09 06-Chp-IPD-Analysis.Rmd
Range of small fixes (randomised, ApoE, etc)

Previous work has suggested a difference in the effect of lipids on dementia risk based on participant age and sex,[@ancelin2013; @mielke2010] though the systematic review presented in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading) did not identify strong evidence for an effect of sex or age at lipid measurement on the relationship between lipids and dementia. However, the number of included studies in the lipid fraction meta-analyses was small, due to both the relatively small number of studies reporting on the lipids exposure and the poor reporting of summary statistics of participant characteristics in those that did. Additionally, best practice guidance recommends against basing the decision to perform an IPD meta-analysis on the between-study heterogeneity in a summary data meta-analysis, as similar distributions of participant covariates across studies may mask a true effect.[@riley2020]

@@

The first approach focused on previously analysed observational prospective cohort studies examining the effect of blood lipid levels on dementia outcomes, as identified by the systematic review presented in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading). The data sources used in each of these cohort analyses were screened against the criteria listed in the previous section, and eligible cohorts were approached for data access. In the first instance, the first/corresponding author of the publication was emailed (see Appendix \@ref(ipd-email-collab) for a copy of the correspondence sent to potential collaborators) in Autumn 2020. If this approach did not elicit a response within two months, the last author was contacted, on the basis that the first/corresponding author may have been a more junior member of the research group who had moved to a different institution.

@@

A range of additional variables were included in the analysis, intended to address the potential for confounding. With an awareness that discrepancies in available covariates is a common issue in IPD studies, I defined an idealised set of covariate domains to be age, sex, education, BMI, _ApoE4_ status, smoking/alcohol status, ethnicity, and prevalent cardiovascular disease. This set covers key risk factors for dementia/Alzheimer's disease in addition to general cardiovascular risk factors. Details on how these variables were coded, given the available data, are presented in Section \@ref(ipd-covar-definition).

@@

With respect to the application to DPUK cohorts, where a dedicated project manager liaises with data owners on the applicants' behalf, the overall response rate was higher. However, even using this streamlined approach, a positive response was obtained for approximately half (N=9, 53%) of the approached cohorts.

@@

The Caerphilly Prospective Study (CaPS) is a longitudinal study of men in South Wales, UK.[@zotero-15398;@elwood2013] Blood lipids (TC, LDL-c, HDL-c and TG) were measured at baseline in 1979-1983, and from Phase III (1989-1993) onwards, a battery of cognitive tests were introduced. The cohort contains data on dementia outcomes sub-classified as vascular and non-vascular dementia. Data was available for all covariates of interest except for education, BMI and _ApoE4_ status.

@@

The European Prospective Investigation of Cancer (EPIC) - Norfolk is a population-based cohort, containing men and women recruited from 35 general practices in Norfolk between 1993 and 1998.[@riboli1997; @riboli2002] The added evidential value of the EPIC cohort is small, given the the fact that the data obtained contains only 8 dementia events. The cohort contains no information on dementia subtype, while all covariates of interest bar _ApoE4_ were available.

@@

The Whitehall II study is a prospective cohort study of men and women recruited between 1985 and 1989 from the civil service in London.[@marmot2005] The cohort is linked with the Hospital Episode Statistics (HES) database, a database containing details of participant events at NHS hospitals in England, which was used to define covariates at baseline and dementia outcomes.[@zotero-15403] Information was available on dementia type, which was classified as all-cause dementia, Alzheimer's disease or vascular dementia. This data source contained details on all covariates of interest except for _ApoE4_ status.

Of note, the Whitehall II cohort was analysed in one of the included studies identified by the systematic review presented in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading),[@tynkkynen2018] meaning that a comparison between the published result and the analysis reported here was was possible.

@@

Missing values in collected variables was common across the cohorts. A matrix of covariates, and whether they were not available or contained missing values for each included cohort, can be seen Appendix \@ref(appendix-ipd-covariate-matrix). The Whitehall II and EPIC cohorts contained data on all but one covariates of interest, while three missing covariates were identified in the CaPS cohort, namely education, BMI and ethnicity. No included cohort provided information on the _ApoE4_ status of participants, and so this variable could not be adjusted for in the analysis.

@@

The results from the main effect analysis across the varying lipid fractions on each dementia,  considered can be seen in Figures \@ref(fig:mainEffectDem) & \@ref(fig:mainEffectVad), respectively. There was very weak evidence for an association of any lipid level with either all-cause or vascular dementia, with the exception of a harmful association between raised triglycerides and vascular dementia (`r ipd_vasdem`).

@@

This analysis requested data from `r n_applied$total` data sources, but only obtained data from three, all of which were based in the United Kingdom. No evidence for an effect of lipids on the risk of dementia or related outcomes was identified, except for a harmful association of raised triglycerides with risk of vascular dementia. Similarly, there was very weak evidence for an interaction of the effect of lipids levels on dementia outcomes with participants' age, grouped into 5 year bands, or sex.

A detailed comparison of the findings presented above with the existing evidence base identified by the systematic review (Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading)), is presented as part of the triangulation exercise in Chapter \@ref(tri-heading). As such, this discussion will not provide a detailed comparison of the results of this analysis with other published literature, except to compare between this analysis and previously published results using the same data source.

@@

A key limitation of this analysis is the potential for uncontrolled confounding. Across all cohorts, adjustment for _ApoE4_ was not possible as I did not have access to genetic data on participants. _ApoE4_ is a strong risk factor for both increased LDL-c levels and Alzheimer's disease,[@bennet2007;@safieh2019] and failing to adjust for this factor means residual confounding in the LDL-c fraction results is likely.

@@

Investigation of this discrepancy indicated that it was likely to be due to the Whitehall snapshot that each analysis had access to, as comparison of summary statistics indicated that this analysis had substantially more dementia events (n=287 in this analysis vs. n=114 in Tynkkynen _et al._). The covariates adjusted for in each analysis were similar, though the previous analysis had access to genetic data, allowing it to adjust for _ApoE4_ status. However, given that _ApoE4_ is a risk factor for increased LDL-c rather than triglycerides,[@bennet2007] it seems unlikely that additional adjustment for this variable is responsible for the discrepancy in findings observed.

@@

Finally, this analysis provides new evidence on a previously unexplored outcome (vascular dementia), adding to the extremely small evidence base for this outcome identified by the systematic review presented in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading). This is relevant even in the previously analysed WhiteHall II cohort, as the previous analysis by Tynkkynen _et al._ did not report on this outcome.

@@

- I found very weak evidence for an association of any lipid with either all-cause or vascular dementia, except for an increased risk of vascular dementia associated with raised triglycerides. Similarly, there was very weak evidence that the association of blood lipids and dementia outcomes varied by participant age or sex.

>>>>

Previous work has suggested a difference in the effect of lipids on dementia risk based on participant age and sex,[@ancelin2013; @mielke2010] though the systematic review presented in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading) did not identify strong evidence for an effect of sex or age at lipid measurement on the relationship between lipids and dementia. However, the number of included studies in the lipid fraction meta-analyses was small, due to both the relatively small number of studies reporting on the lipids exposure and the poor reporting of summary statistics of participant characteristics in those that did. Additionally, best practice guidance recommends against basing the decision to perform an IPD meta-analysis on the between-study heterogeneity in a summary data meta-analysis, as similar distributions of participant covariates across studies may mask a true effect.[@riley2020]

@@

The first approach focused on previously analysed observational prospective cohort studies examining the effect of blood lipid levels on dementia outcomes, as identified by the systematic review presented in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading). The data sources used in each of these cohort analyses were screened against the criteria listed in the previous section, and eligible cohorts were approached for data access. In the first instance, the first/corresponding author of the publication was emailed (see Appendix \@ref(ipd-email-collab) for a copy of the correspondence sent to potential collaborators) in Autumn 2020. If this approach did not elicit a response within two months, the last author was contacted, on the basis that the first/corresponding author may have been a more junior member of the research group who had moved to a different institution.

@@

A range of additional variables were included in the analysis, intended to address the potential for confounding. With an awareness that discrepancies in available covariates is a common issue in IPD studies, I defined an idealised set of covariate domains to be age, sex, education, BMI, _Apo_$\mathcal{E}4$ status, smoking/alcohol status, ethnicity, and prevalent cardiovascular disease. This set covers key risk factors for dementia/Alzheimer's disease in addition to general cardiovascular risk factors. Details on how these variables were coded, given the available data, are presented in Section \@ref(ipd-covar-definition).

@@

With respect to the application to DPUK cohorts, where a dedicated project manager liaises with data owners on the applicants' behalf, the overall response rate was higher. However, even using this streamlined approach, a positive response was obtained for approximately half (N =9, 53%) of the approached cohorts.

@@

The Caerphilly Prospective Study (CaPS) is a longitudinal study of men in South Wales, UK.[@zotero-15398;@elwood2013] Blood lipids (TC, LDL-c, HDL-c and TG) were measured at baseline in 1979-1983, and from Phase III (1989-1993) onwards, a battery of cognitive tests were introduced. A dementia was derived using data obtained during Phase V (2002-2004), giving between 19-25 years of follow-up. The cohort contains data on dementia outcomes sub-classified as vascular and non-vascular dementia. Data was available for all covariates of interest except for education, BMI and _Apo_$\mathcal{E}4$ status.

@@

The European Prospective Investigation of Cancer (EPIC) - Norfolk is a population-based cohort, containing men and women recruited from 35 general practices in Norfolk between 1993 and 1998.[@riboli1997; @riboli2002] Data provided by this cohort for the analysis, dementia was ascertained at at the 5th Health Checkup (2016-2018), giving between 18-25 years of follow-up. The added evidential value of the EPIC cohort is small, given the the fact that the data obtained contains only 8 dementia events. The cohort contains no information on dementia subtype, while all covariates of interest bar _Apo_$\mathcal{E}4$ were available.

@@

The Whitehall II study is a prospective cohort study of men and women recruited between 1985 and 1989 from the civil service in London.[@marmot2005] Lipids measurements were available from the third wave of the study, conducted between 1991-1993. The cohort is linked with the Hospital Episode Statistics (HES) database, a database containing details of participant events at NHS hospitals in England, which was used to define covariates at baseline and dementia outcomes.[@zotero-15403] HES data for dementia outcomes was avaialle up to March 2015, providing between 22-24 years of follow-up. Information was available on dementia type, which was classified as all-cause dementia, Alzheimer's disease or vascular dementia. This data source contained details on all covariates of interest except for _Apo_$\mathcal{E}4$ status.

Of note, the Whitehall II cohort was analysed in one of the included studies identified by the systematic review presented in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading),[@tynkkynen2018] meaning that a comparison between the published result and the analysis reported here was was possible.

@@

Missing values in collected variables was common across the cohorts. A matrix of covariates, and whether they were not available or contained missing values for each included cohort, can be seen Appendix \@ref(appendix-ipd-covariate-matrix). The Whitehall II and EPIC cohorts contained data on all but one covariates of interest, while three missing covariates were identified in the CaPS cohort, namely education, BMI and ethnicity. No included cohort provided information on the _Apo_$\mathcal{E}4$ status of participants, and so this variable could not be adjusted for in the analysis.

@@

The results from the main effect analysis across the varying lipid fractions on each dementia,  considered can be seen in Figures \@ref(fig:mainEffectDem) & \@ref(fig:mainEffectVad), respectively. There was weak evidence for an association of any lipid level with either all-cause or vascular dementia, with the exception of a harmful association between raised triglycerides and vascular dementia (`r ipd_vasdem`).

@@

This analysis requested data from `r n_applied$total` data sources, but only obtained data from three, all of which were based in the United Kingdom. No evidence for an effect of lipids on the risk of dementia or related outcomes was identified, except for a harmful association of raised triglycerides with risk of vascular dementia. Similarly, there was weak evidence for an interaction of the effect of lipids levels on dementia outcomes with participants' age, grouped into 5 year bands, or sex.

A detailed comparison of the findings presented above with the existing evidence base identified by the systematic review (Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading)), is presented as part of the triangulation exercise in Chapter \@ref(tri-heading). As such, this discussion will not provide a detailed comparison of the results of this analysis with other published literature, except to compare between this analysis and previously published results using the same data source.

@@

A key limitation of this analysis is the potential for uncontrolled confounding. Across all cohorts, adjustment for _Apo_$\mathcal{E}4$ was not possible as I did not have access to genetic data on participants. _Apo_$\mathcal{E}4$ is a strong risk factor for both increased LDL-c levels and Alzheimer's disease,[@bennet2007;@safieh2019] and failing to adjust for this factor means residual confounding in the LDL-c fraction results is likely.

@@

Investigation of this discrepancy indicated that it was likely to be due to the Whitehall snapshot that each analysis had access to, as comparison of summary statistics indicated that this analysis had substantially more dementia events (N =287 in this analysis vs. n=114 in Tynkkynen _et al._). The covariates adjusted for in each analysis were similar, though the previous analysis had access to genetic data, allowing it to adjust for _Apo_$\mathcal{E}4$ status. However, given that _Apo_$\mathcal{E}4$ is a risk factor for increased LDL-c rather than triglycerides,[@bennet2007] it seems unlikely that additional adjustment for this variable is responsible for the discrepancy in findings observed.

@@

Finally, this analysis provides new evidence on a previously unexplored outcome (vascular dementia), adding to the extremely small evidence base for this outcome identified by the systematic review presented in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading). This is relevant even in the previously analysed WhiteHall II cohort, as the previous analysis by Tynkkynen _et al._ did not report on this outcome.

@@

- I found weak evidence for an association of any lipid with either all-cause or vascular dementia, except for an increased risk of vascular dementia associated with raised triglycerides. Similarly, there was weak evidence that the association of blood lipids and dementia outcomes varied by participant age or sex.

### 2021-12-09 07-Chp-Triangulation.Rmd
Range of small fixes (randomised, ApoE, etc)

Analysis of lipid levels across both the systematic review (Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading)) and the IPD analysis (Chapter \@ref(ipd-heading)) both found extremely weak evidence for an association of any lipid fraction with all-cause dementia (TC `r obsLipids$Dementia_TC$estimate`; LDL-c `r obsLipids$Dementia_LDL$estimate`; HDL-c `r obsLipids$Dementia_HDL$estimate`; triglycerides `r obsLipids$Dementia_TG$estimate`). Similarly, meta-analysis of studies identified by the review which examined a binary hypercholesterolemia provided weak evidence for an association of this risk factor with Alzheimer's disease (`r obsHyperchol$Dementia$estimate`). Finally, a meta-analysis of RCTs of statin use (`r rct_statin_acd`) and of MR analyses of genetic instruments that mimic statin use (`r mrStatin$Dementia$estimate`) both found weak evidence of an association, providing indirect evidence on the effect of lipid lowering on this outcome.

@@

@@

As an illustrative example of the process using a single result, the bias and indirectness inherent to the estimate of the effect of statin use on Alzheimer's disease in the CPRD (as analysed in Chapter \@ref(cprd-analysis-heading)) are considered here. For this result, there were three domains at greater than low risk of bias: Bias due to confounding, Bias due to definition of the outcome, and Bias in the selection of the reported result (Figure \@ref(fig:biasDirectionSingle)). In Domain 1 (bias due to confounding), the study did not adjust for _ApoE4_ status, which is predicted to mask any true association between statins and Alzheimer's disease (see Section \@ref(cprd-limitations) for a fuller discussion). As such, the bias is predicted to _favour the comparator_ and so the absolute direction of bias is to the right. In Domain 7 (bias in measurement of the outcome), there was a high risk of differential misclassification on the basis of the exposure. Here, history of statin use may make a diagnosis of vascular dementia more likely than Alzheimer's disease, and so for this outcome, this bias is predicted to _favour the experimental_ (in this case, an intervention). As such, the absolute direction of bias is to the left. In Domain 7 (bias in selection of the reported result), for consistency with the other NRSI, in the absence of a protocol the bias is assumed to be _away from the null_. Given the position of the effect estimate below the null, the absolute direction of bias in this domain is to the left.

@@

From the data sources identified in Section \@ref(tri-data-sources), I identified 18 results relevant to my first causal question of interest. Most (n=17) were identified via the systematic review,`r ldl_ad_citations` while Chapter \@ref(cprd-analysis-heading) provided an additional estimate from the CPRD analysis.  The extent, type and direction of predicted bias for each result can be seen in the bias-direction plot presented in Figure \@ref(fig:ldlAdBiasDirection), stratified by study design.

@@

@@

@@

Ideally, these prior distributions would be based on empirical data on the effect of different domains of bias/indirectness on a result. Most meta-epidemiological studies examining the effect of different methodological issues studied only randomised controlled trials,[@amer2021; @page2016] and the estimates of bias they produce may not hold for non-randomised study designs. However, the prior distributions defined here are broadly comparable to previously reported estimates of the impact of bias in non-randomised trials. For example, a previous simulation study estimated that, in the absence of true effect of LDL-c, _ApoE4_ would induce a RR of 1.09 with risk of dementia per 1-SD increase in LDL-c.[@iwagami2021] This is comparable to the mean assigned to the prior distribution mapped to "Moderate" additive bias (N(0.08,0.05), Table \@ref(tab:priorsAdd-table)).

>>>>

Analysis of lipid levels across both the systematic review (Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading)) and the IPD analysis (Chapter \@ref(ipd-heading)) both found extremely weak evidence for an association of any lipid fraction with all-cause dementia (TC `r obsLipids$Dementia_TC$estimate`; LDL-c `r obsLipids$Dementia_LDL$estimate`; HDL-c `r obsLipids$Dementia_HDL$estimate`; triglycerides `r obsLipids$Dementia_TG$estimate`). Similarly, meta-analysis of studies identified by the review which examined a binary hypercholesterolemia provided weak evidence for an association of this risk factor with Alzheimer's disease (`r obsHyperchol$Dementia$estimate`). Finally, a meta-analysis of RCTs of statin use (`r rct_statin_acd`) and of MR analyses of genetic instruments that mimic statin use (`r mrStatin$Dementia$estimate`) both found weak evidence of an association, providing indirect evidence on the effect of lipid lowering on this outcome.

@@

<!-- TODO The results should be standarised before calculating direction of effect, because other-wise it is pulling it in the wrong direction -->

@@

As an illustrative example of the process using a single result, the bias and indirectness inherent to the estimate of the effect of statin use on Alzheimer's disease in the CPRD (as analysed in Chapter \@ref(cprd-analysis-heading)) are considered here. For this result, there were three domains at greater than low risk of bias: Bias due to confounding, Bias due to definition of the outcome, and Bias in the selection of the reported result (Figure \@ref(fig:biasDirectionSingle)). In Domain 1 (bias due to confounding), the study did not adjust for _Apo_$\mathcal{E}4$ status, which is predicted to mask any true association between statins and Alzheimer's disease (see Section \@ref(cprd-limitations) for a fuller discussion). As such, the bias is predicted to _favour the comparator_ and so the absolute direction of bias is to the right. In Domain 7 (bias in measurement of the outcome), there was a high risk of differential misclassification on the basis of the exposure. Here, history of statin use may make a diagnosis of vascular dementia more likely than Alzheimer's disease, and so for this outcome, this bias is predicted to _favour the experimental_ (in this case, an intervention). As such, the absolute direction of bias is to the left. In Domain 7 (bias in selection of the reported result), for consistency with the other NRSI, in the absence of a protocol the bias is assumed to be _away from the null_. Given the position of the effect estimate below the null, the absolute direction of bias in this domain is to the left.

@@

From the data sources identified in Section \@ref(tri-data-sources), I identified 18 results relevant to my first causal question of interest. Most (N =17) were identified via the systematic review,`r ldl_ad_citations` while Chapter \@ref(cprd-analysis-heading) provided an additional estimate from the CPRD analysis.  The extent, type and direction of predicted bias for each result can be seen in the bias-direction plot presented in Figure \@ref(fig:ldlAdBiasDirection), stratified by study design.

@@

<!-- TODO This needs to have the heterogeneity values too -->

@@

<!-- TODO There is a very slight effect in the unadjusted here, but reduces to blah blah on adjustmet -->

@@

Ideally, these prior distributions would be based on empirical data on the effect of different domains of bias/indirectness on a result. Most meta-epidemiological studies examining the effect of different methodological issues studied only randomised controlled trials,[@amer2021; @page2016] and the estimates of bias they produce may not hold for non-randomised study designs. However, the prior distributions defined here are broadly comparable to previously reported estimates of the impact of bias in non-randomised trials. For example, a previous simulation study estimated that, in the absence of true effect of LDL-c, _Apo_$\mathcal{E}4$ would induce a RR of 1.09 with risk of dementia per 1-SD increase in LDL-c.[@iwagami2021] This is comparable to the mean assigned to the prior log-normal distribution mapped to "Moderate" additive bias ($N(0.09,0.05)$, Table \@ref(tab:priorsAdd-table)).

### 2021-12-09 08-Discussion.Rmd
Range of small fixes (randomised, ApoE, etc)

The first is a large-scale analysis of 953,635 patients in the CPRD, followed from their first LDL-c measurement.[@iwagami2021] The study found a slight increased risk of dementia was associated higher LDL-c measured at mid-life, driven by the finding for the Alzheimer's disease subgroup. These results are consistent with the findings presented in this thesis, particularly given the lack of adjustment for confounding by ApoE4, which was estimated by the study to induce a RR of 1.09 per 1-SD increase in LDL-c in the absence of a true causal effect.

The second analysis examined 502,226 participants in the UK Biobank, a large population-based prospective cohort study.[@gong2021; @sudlow2015] The analysis found no association between lipids measured at mid-life and dementia outcomes, comparable to the results of this thesis. However, similar to the study discussed above, this analysis did not have access to genetic data and so may be subject to residual confounding by _ApoE4_ genotype.

@@

Thirdly, missing evidence was a common limitation across this thesis. In Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading), several studies stated that they did not report the results of an analysis because the results were not significant. This was particularly common among non-randomised studies of exposure examining blood lipids directly. Similarly, an absence of evidence on vascular dementia, potentially due to a publication bias mechanism, limited the analysis of this outcome. Missing evidence was also as an issue in the IPD meta-analysis reported in Chapter \@ref(ipd-heading), in the form of a poor response to data access requests, while the absence of empirically-based distributions of bias/indirectness for use with the triangulation framework limited the credibility of the results produced.

>>>>

The first is a large-scale analysis of 953,635 patients in the CPRD, followed from their first LDL-c measurement.[@iwagami2021] The study found a slight increased risk of dementia was associated higher LDL-c measured at mid-life, driven by the finding for the Alzheimer's disease subgroup. These results are consistent with the findings presented in this thesis, particularly given the lack of adjustment for confounding by _Apo_$\mathcal{E}4$, which was estimated by the study to induce a RR of 1.09 per 1-SD increase in LDL-c in the absence of a true causal effect.

The second analysis examined 502,226 participants in the UK Biobank, a large population-based prospective cohort study.[@gong2021; @sudlow2015] The analysis found no association between lipids measured at mid-life and dementia outcomes, comparable to the results of this thesis. However, similar to the study discussed above, this analysis did not have access to genetic data and so may be subject to residual confounding by _Apo_$\mathcal{E}4$ genotype.

@@

Thirdly, missing evidence was a common limitation across this thesis. In Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading), several studies stated that they did not report the results of an analysis because the results were not significant. This was particularly common among non-randomised studies of exposure examining blood lipids directly. Similarly, an absence of evidence on vascular dementia, potentially due to a publication bias mechanism, limited the analysis of this outcome. Missing evidence was also as an issue in the IPD meta-analysis reported in Chapter \@ref(ipd-heading), in the form of a poor response to data access requests, while the absence of empirically-based distributions of bias/indirectness for use with the triangulation framework limited the credibility of the results produced.
